---
title: "Reproducible Research Assignment1"
output: html_document
---
Loading and preprocessing the data
```{r}
readActivity = read.csv("/DataScience/R-Files/RepResearch/activity.csv",head=TRUE, sep=",")
```

Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
filterData <- na.omit(readActivity)
totSteps <- aggregate(filterData$steps, by=list(filterData$date), FUN=sum)
names(totSteps) <- c("Date","Steps")
```

What is mean total number of steps taken per day?

Make a histogram of the total number of steps taken each day
```{r}
hist(totSteps$Steps)
```

Calculate and report the mean and median total number of steps taken per day
```{r}
mean(totSteps$Steps)
median(totSteps$Steps)
```
What is the average daily activity pattern?

Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and 

the average number of steps taken, averaged across all days (y-axis)
```{r}
AvgInterval <- aggregate(filterData$steps, by=list(filterData$interval), FUN=mean)
names(AvgInterval) <- c("Interval","Steps")
plot(AvgInterval$Steps ~ AvgInterval$Interval, type="l")
```

Which 5-minute interval, on average across all the days in the dataset, 
contains the maximum number of steps?
```{r}
AvgInterval[which.max(AvgInterval$Steps), ]$Interval
```

Imputing missing values
Note that there are a number of days/intervals where there are missing values (coded as NA). 
The presence of missing days may introduce bias into some calculations or summaries of the data.
Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(is.na(readActivity$steps))
```
Devise a strategy for filling in all of the missing values in the dataset. 
The strategy does not need to be sophisticated. For example, you could use the mean/median 
for that day, or the mean for that 5-minute interval, etc.
Create strategy for filing in all of the missing values with the mean for that interval.
```{r}
getAvgSteps <- function(interval) 
         { 
          AvgInterval[AvgInterval$Interval == interval, ]$Steps
} 
```
Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
activityUpd <- readActivity
count <- 0  
for (i in 1:nrow(activityUpd)) 
    {
     if (is.na(activityUpd[i, ]$steps)) {
        activityUpd[i, ]$steps <- getAvgSteps(activityUpd[i, ]$interval)
        count <- count + 1
    }
}
cat("Total ", count, "NA values were filled.")
```

Make a histogram of the total number of steps taken each day and Calculate and 
report the mean and median total number of steps taken per day. Do these values differ 
from the estimates from the first part of the assignment? What is the impact of 
imputing missing data on the estimates of the total daily number of steps?

```{r}
totSteps2 <- aggregate(activityUpd$steps, by=list(activityUpd$date), FUN=sum)
names(totSteps2) <- c("Date","Steps")
hist(totSteps2$Steps)
```

Calculate and report the mean and median total number of steps taken per day
```{r}
mean(totSteps2$Steps)
median(totSteps2$Steps)
```
Do these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total daily number of steps?
 

The value of mean is same for before and after imputing missing data.

There is a little difference for median value before and after imputing missing data.


Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. 
Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels - "weekday" 
and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activityUpd$day <- ifelse(as.POSIXlt(as.Date(activityUpd$date))$wday%%6 == 0, "weekend", "weekday" )
```
 weekday for Monday to Friday and weekend for Saturday and Sunday
```{r}
activityUpd$day = factor(activityUpd$day, levels = c("weekday", "weekend"))
```
Make a panel plot containing a time series plot (i.e. type = "l") 
of the 5-minute interval (x-axis) and the average number of steps taken, averaged across 
all weekday days or weekend days (y-axis). 
```{r}
AvgInterval3 <- aggregate(steps ~ interval + day, activityUpd, mean)
library(lattice)
xyplot(steps ~ interval | factor(day), data = AvgInterval3, type = "l", aspect = 1/2 )
